### 2023-09-02 chapters/background.tex
Some formatting and changes to labels
Write on introduction

In society, the concept of uncertainty goes by many names, and its meaning can vary depending on the specific context. However, across most quantitative scientific fields, a consensus definition appears to align with the following \cite{hubbard_how_2014}:
%
\begin{center}
    \textit{Uncertainty is the lack of certainty; a state of limited knowledge where it is impossible to exactly describe the existing state, a future outcome, or more than one possible outcome.}
\end{center}
%
While such a purely lexical definition of uncertainty might prompt a philosophical inquiry, it highlights an important connection between uncertainty, information and probability. When information is not complete, we can describe the state of the world only with uncertainty; we are simply not certain. In that sense, uncertainty is a word we use to refer to information that is missing. A natural way to describe a state of the world with uncertainty is to use probabilities. To our luck, information theory provides exactly that; a mathematically rigorous method for quantifying information through the language of probabilities. In this context, uncertainty can be understood as the information entropy of a random variable \cite{mackay_information_2003}.

@@

\lesstodo[inline]{Come up with a suitable example explaining the intuition behind information entropy.}

>>>>

In society, the concept of uncertainty goes by many names, and its meaning can vary depending on the specific context. However, across most quantitative scientific fields, a consensus definition appears to align with the following \cite{hubbard_how_2014}:
%
\begin{center}
    \textit{Uncertainty is the lack of certainty; a state of limited knowledge where it is impossible to exactly describe the existing state, a future outcome, or more than one possible outcome.}
\end{center}
%
Although such a purely lexical definition of uncertainty might prompt a philosophical inquiry, it highlights an important connection between uncertainty, information and probability. When information is limited, we can describe the state of the world only with uncertainty; we are simply not certain. In that sense, uncertainty is a word we use to refer to the information that is missing, whether we know what is missing or not. 
A natural way to describe an uncertain world state is to use probabilities. To our luck, information theory provides exactly that; a mathematically rigorous method for quantifying information through the language of probability theory. In the following, we shall see that, in this context, uncertainty can be understood as the information entropy of a random variable \cite{mackay_information_2003}.

@@

\todo[inline]{Come up with a suitable example explaining the intuition behind information entropy.}

### 2023-09-02 chapters/introduction.tex
Write on introduction

\chapter[introduction]{Introduction}\label{chp:introduction}

% ~9 pages

% Information overlead + administrative burden + lack of resources + rapid scientific progress in medicine and healthcare
In an era where technology increasingly intertwines with healthcare, artificial intelligence is on the verge of becoming part of standard medical practice. 
Growing administrative burdens, rapid scientific progress in medicine, and aging populations have created a need to rethink how medical professionals are best enabled to successfully do their job. 
The application of artificial intelligence in medical decision-making, particularly in diagnosing and recommending treatments, is likely to hold significant potential for improving patient outcomes by reducing the effect of these issues. 
However, this prospect also raises concerns related to the reliability of such models as their decisions impact critical aspects of human health and well-being.

@@

However, these new computing tools also brought new challenges. The Babylonian system of mathematics did not formalize the difference between rational and irrational numbers often leading to approximation errors and the common practice of approximating $\pi$ with 3. Apart from limitations inherent to the tools, inadequate human proficiency in their use was also a source of error. The Plimpton 322 clay tablet, for instance, contains several errors presumably made by a novice in Babylonian mathematics (\cref{fig:plimpton_332}) \cite{britton_plimpton_2011}.
These examples highlight the importance of understanding the limitations of technology, even in ancient times.

\begin{figure}[t]
    \centering
    \includegraphics[width=0.80\textwidth]{plimpton_322.png}
    \caption{Technology is difficult. The scribe of the Plimpton 322 clay tablet ($\sim$1800 BCE) made several errors when listing Pythagorean triplets, presumably struggling to grasp Babylonian mathematics.}
    \label{fig:plimpton_332}
\end{figure}

While the quest to ease human decision-making has been ongoing for thousands of years, it was not until 1642 that the first mechanical calculator was successfully built by Blaise Pascal. And while the idea of a programmable computer with applications beyond pure arithmetic was first proposed by Ada Lovelace in 1843, 
it was not until the 1940s and 50s that % / 
% it took one-hundred years before / 
the first programmable, electronic, general-purpose digital computers were developed, starting with the ENIAC in 1945 (\cref{fig:eniac_programmers}) \cite{georges_universal_2001}. 
As they became more accessible, general-purpose computers were adopted widely across society for various applications in science and industry, including healthcare.
The societal changes brought about by the widespread adoption of computer technology through the latter half of the 20th century were so profound that some scholars refer to the current era as the information age \cite{georges_universal_2001, harari_sapiens_2011}.
\todo[inline]{Expand on the history of modern computing.}

\begin{figure}[t]
    \centering
    \includegraphics[width=0.95\textwidth]{eniac_programmers.jpeg}
    \caption{Ruth Lichterman (left) and Marlyn Wescoff (middle) were two of the female programmers of the ENIAC. Photo: \copyright Corbis/ Getty Images.}
    \label{fig:eniac_programmers}
\end{figure}

With the advent of the internet and the widespread adoption of digital technologies, the amount of data available to decision-makers has grown exponentially. 
This has led to a surge of interest in the scientific field of machine learning, a subfield of artificial intelligence that focus on developing algorithms that can learn from data and make predictions to guide decisions based on that knowledge. 
Such methods have set new standards for the performance of computer systems in various domains, including speech recognition, computer vision, and natural language processing \cite{lecun_deep_2015}.
In recent years, machine learning has also been applied within healthcare, where it has shown promising results in tasks such as medical imaging \cite{lundervold_overview_2019}, drug discovery \cite{chen_rise_2018}, and clinical decision support \cite{cite15, cite14}.

However, these technologies come with the same risks as previous generations did, and in some cases, they may exacerbate them. Inherent limitations of the technology itself, human error, and the complexity of the real world are all factors that can lead to failures. A pertinent example is the use of machine learning in autonomous driving systems: 
In 2018, the first incident of a self-driving car killing a pedestrian took place in Tempe, Arizona. 
The car, an SUV operated by Uber, was driving in autonomous mode when it struck a pedestrian crossing the street, pushing her bicycle. 
According to the investigation by the National Transportation Safety Board, the car's sensors detected the pedestrian $5.6$ seconds before the crash, but it did not correctly predict her path or reduce the SUV's speed (\cref{fig:uber_nhsa_accident}). 
Specifically, during those seconds, the system incorrectly classified the pedestrian more than ten times, first as a vehicle, then as an unknown object and ultimately as a bicyclist, each time changing, or resetting, the predicted path. 
About one second before the crash, the system determined that a collision was imminent, but the situation exceeded the constraints within which the autonomous driving system was allowed to operate, and the car's safety driver failed to intervene \cite{nationaltransportationsafetyboardnhsa_collision_2019}.

@@

Though this incident is an extreme one, it highlights the challenges of applying machine learning in real-world scenarios and underlines the need for machine learning systems to be able to accurately assess their own uncertainty and operate accordingly. 
Recently, European policymakers proposed a set of requirements for applications of machine learning to ensure that they are safe, reliable, and trustworthy \cite{europeancommission_briefing_2021}. This AI Act groups applications into three categories based on the potential risk they pose to society: limited risk, high risk, and unacceptable risk. Limited risk applications, for instance within gaming, must meet transparency requirements, and unacceptable risk applications, such as public use of facial recognition, are banned. Applications within healthcare are considered high risk and must meet a strict set of requirements focusing on anti-discrimination and robustness. A recent study by the European Parliamentary Research Services concluded: \textit{"Future AI solutions for healthcare should be implemented by integrating uncertainty estimation, a relatively new field of research that aims to provide clinicians with clinically useful indications on the degree of confidence in AI predictions"} \cite{europeanparliament_artificial_2022}.

@@

\noindent The research presented in this thesis will focus on the use of machine learning in relation to uncertainty estimation, speech processing, and medical conversations. 
The research project of this thesis was formulated and carried out in collaboration with the Danish company Corti, which develops computer software for medical decision support. 
In the rest of this chapter, we will provide a high-level introduction to these topics and discuss the motivation behind the research project.

\section{Motivation and Corti's use-cases}

At the start of this project in 2020, Corti's main product was a software system providing decision-support for emergency medical dispatchers.
The system 
% was in use in several countries, including the 1-1-2 line in Copenhagen, Denmark and the 9-1-1 line in Seattle, Washington in the United States and 
provides an interface for navigating a graph-based protocol, defined by the local emergency medical services, which dispatchers use to triage patients during emergency calls. Calls are transcribed in real-time using customer-domain automatic speech recognition and the system makes suggestions for the dispatcher to use in the conversation with the caller, including notifications about forgotten protocol questions \cite{havtorn_multiqt_2020} and potential cases of cardiac arrests \cite{cite15, cite14}. The system logs the details about the call and the dispatcher's actions to be used for later review and quality assurance.

\subsection{Stroke recognition in emergency calls}
In 2021, Corti entered a research collaboration with the Capital Region of Denmark and the Copenhagen Emergency Services to develop a system for recognizing stroke cases in calls to the 1-1-2 emergency line and the 1813 medical helpline. 
Stroke is one of the biggest causes of disability and death worldwide \cite{cite1,cite2,cite3}. Effective treatment is highly time-sensitive \cite{cite4,cite5} and the most common gateway to specialized treatment and hospital admittance is through prehospital telehealth services like emergency medical call centers, nurse advice call lines, and out-of-hours health services \cite{cite6,cite7}. However, studies have found that approximately half of all patients with stroke do not receive the correct triage for their condition from call-takers \cite{cite10,cite11,cite12}. This may be due to the fact that stroke is a complex condition with a wide range of symptoms and that the symptoms may be difficult to recognize over the phone. Additionally, out of total call volume, stroke cases are relatively rare, occurring in only $0.25\%$ of all calls to the Copenhagen 1-1-2 emergency line in 2021, which makes it difficult for call-takers to gain practical experience in recognizing stroke cases.

A reasonable machine learning pipeline for recognizing stroke might consist of an automatic speech recognition model to transcribe the conversation and a binary classifier to estimate the probability that what the transcript describes is a case of stroke. 
% The performance of such a system would depend on the quality of the automatic speech recognition model, the quality of the classifier, and the quality of the data used to train them. 
In such a system, many factors can lead to classification errors, and when that happens, we would like the model to express high degree of uncertainty. 
The automatic speech recognition model might be make an error in the transcription of particular word or phrase due to a noisy environment, overlapping, slurred or mumbling speech, or words outside the training distribution of the model. 
The classifier might misclassify a conversation due to ambiguous symptoms, a general lack of information, or errors of the speech recognizer propagated via the transcription. 
How to best make such models accurately represent and estimate the uncertainty of given predictions, and how to use it to improve the value of such a system, especially in cooperation with the call taker is an open question.
What is clear, however, is that the ability of such a system to accurately estimate the uncertainty of given predictions is crucial for its safe and reliable deployment in the real world. Overconfident predictions could lead to misdiagnoses and potentially fatal consequences for patients. Underconfident predictions could lead to unnecessary delays in treatment and increased costs for the healthcare system.

% Corti's machine learning systems are not exempt from the challenges that come with real-world applications. The development is an ongoing effort to improve performance and to ensure robustness, reliability and compliance with current regulations. This thesis

% One of the challenges faced by Corti's systems is the need to be able to recognize and handle rare events.
% For instance, the system for emergency medical dispatchers is designed to recognize cases of cardiac arrest in emergency calls.
% Cardiac arrest is a rare event, occurring in only $0.1\%$ of emergency calls \cite{cite14}, and the system must be able to recognize it in order to provide the dispatcher with the correct protocol to follow.
% The system is trained on a dataset of calls where the dispatcher has indicated that the patient was in cardiac arrest, but the dataset is small and imbalanced, making it difficult to train a model that can reliably recognize cardiac arrest cases.
% In addition, the system must be able to recognize cardiac arrest cases in calls from different countries, where the protocols and the way dispatchers handle calls may differ.
% This is a challenging task as the system must be able to generalize to unseen data and adapt to different contexts.

\subsection{Medical coding of clinical notes}

During the course of the project presented in this thesis, Corti's product portfolio has expanded to include several new products. 
One of these is a software system for medical coding of clinical notes. 
When a patient is admitted to a hospital, the medical staff will write a clinical note describing the patient's diagnosis and the procedures they underwent, including drug prescriptions. 
These notes are then used for billing and reimbursement purposes, as well as for research and quality assurance. 
The clinical notes are written in natural language, usually following a certain structure, and a medical coder will assign a set of medical codes to the note based on its content. 
This process is time-consuming and error-prone task due to the complexity of medical diagnoses and procedures and the immense number of medical codes in the most common standards such as the Internation Classification of Diseases (ICD) versions 10 and 11 with 55,000 and 85,000 codes, respectively \cite{worldhealthorganisationwho_international_2023}.

\todo[inline]{
Discuss the role of uncertainty estimation in medical coding. 
Special challenges include significant class imbalance that cannot be removed by stratification since each example is associated with multiple labels.
Mutually dependent classes also complicates uncertainty estimation. Some classes are mutually exclusive, while others are not. Some classes are easily confused with each other.
}

\section{Machine learning trustworthiness}

Many modern machine learning models are deep neural networks with millions or even billions of parameters. 
These models are often referred to as black-box models as it is difficult to understand how they arrive at their predictions.
The lack of interpretability makes it difficult to assess the reliability of the model's predictions.
Furthermore, a number of factors including the ability of deep neural networks to overfit, or even memorize, the training data \cite{arpit_closer_2017, burg_memorization_2021}, as well as the use of training objectives that are proxies for the evaluation metrics of interest, often lead to models that are miscalibrated; that is, the probability of a class predicted by the model does not reflect the true probability of the predicted class \cite{guo_calibration_2017, kull_temperature_2019}. Commonly, the predicted probabilities are too extreme, leading to overconfident predictions. This means a model will often assign a high probability to the predicted class, even when it is wrong.

\todo[inline]{Expand in this discussion of overconfidence and calibration. E.g.: Calibration is fairly well understood in the context of classification, especially binary, but is harder in structured prediction such as automated speech recognition.}

\section{Scope of the thesis}

The complete set of research published as part of this project covers a varied set of topics. One group of the produced studies is concerned with generative latent variable models and their applications to uncertainty estimation and speech modelling \cite{havtorn_hierarchical_2021,havtorn_benchmarking_2022,bergamin_modelagnostic_2022}. Another group of studies is concerned with self-supervised speech representation learning and automatic speech recognition \cite{borgholt_scaling_2021,borgholt_we_2021,mohamed_selfsupervised_2022,borgholt_brief_2022}. A final group deals with applications within the medical domain including recognition of stroke cases in calls to medical helplines \cite{wenstrup_retrospective_2023} and medical coding of clinical notes \cite{edin_automated_2023}.

@@

% The ancient Egyptians used mathematics for practical purposes like measuring land, constructing buildings, and managing resources. The Rhind Mathematical Papyrus, dated around 1550 BCE, contains various mathematical problems and techniques for calculations, such as multiplication, division, and calculations involving fractions.

% The Babylonians, around 300 BCE, developed a sophisticated system of mathematics, including algorithms for solving quadratic equations and calculating areas of shapes. The Babylonians also used numerical methods for approximating square roots.

% Greek mathematicians like Euclid (circa 300 BCE) formulated mathematical proofs and geometrical concepts that laid the foundation for algorithmic thinking. The "Euclidean algorithm," used to find the greatest common divisor of two numbers, is still taught in schools today.

% Scholars in the Islamic world made significant contributions to mathematics. Al-Khwarizmi (circa 800 CE) wrote a book called "Al-Kitab al-Mukhtasar fi Hisab al-Jabr wal-Muqabala," from which the term "algebra" is derived. The book introduced systematic methods for solving linear and quadratic equations.

>>>>

\chapter[introduction]{Introduction}\label{chp:introduction}
%s
% SIZE: ~9 pages
%
% OUTLINE: Broad introduction to the topic.
% - Motivate the reader.
% - Why should they care about this topic? 
% - What is the historical context of decision-support, what are the practical implications of it, and what happens when it fails.

In an era where technology increasingly intertwines with healthcare, artificial intelligence is on the verge of becoming part of standard medical practice. 
Growing administrative burdens, rapid scientific progress in medicine, and aging populations have created a need to rethink how medical professionals are best enabled to successfully do their job. 
The application of artificial intelligence in medical decision-making, particularly in diagnosing and recommending treatments, is likely to hold significant potential for improving patient outcomes by reducing the effect of these issues. 
However, this prospect also raises concerns related to the reliability of such models as their decisions impact critical aspects of human health and well-being.

@@

However, these new computing tools also brought new challenges. The Babylonian system of mathematics did not formalize the difference between rational and irrational numbers often leading to approximation errors and the common practice of approximating $\pi$ with 3. Apart from limitations inherent to the tools, inadequate human proficiency in their use was also a source of error. The Plimpton 322 clay tablet, for instance, contains several errors presumably made by a novice in Babylonian mathematics (\cref{fig:plimpton_332}) \cite{britton_plimpton_2011}. 
It is easy to forget that the challenges we face today in adopting new technologies are not new but have been faced by humans for thousands of years. Even in ancient times, it was important to understand the limitations of technology and the need for human expertise in using it.

\begin{figure}[t]
    \centering
    \includegraphics[width=0.90\textwidth]{plimpton_322.png}
    \caption{Technology is difficult. The scribe of the Plimpton 322 clay tablet ($\sim$1800 BCE) made several errors when listing Pythagorean triplets, presumably struggling to grapple with Babylonian mathematics.}
    \label{fig:plimpton_332}
\end{figure}

While the quest to ease human decision-making has been ongoing for thousands of years, it was not until 1642 that the first mechanical calculator was successfully built by Blaise Pascal. 
And while the idea of a programmable computer with applications beyond pure arithmetic was first proposed by Ada Lovelace in 1843, 
it was not until the 1940s and 50s that % / 
% it took one-hundred years before / 
the first programmable, electronic, general-purpose digital computers were developed, starting with the ENIAC in 1945 (\cref{fig:eniac_programmers}) \cite{georges_universal_2001}. 
As they became more accessible, general-purpose computers were adopted widely across society for various applications in science and industry, including healthcare.
The societal changes brought about by the widespread adoption of computer technology through the latter half of the 20th century were so profound that some scholars refer to the era as the information age \cite{georges_universal_2001, harari_sapiens_2011}.
\todo[inline]{Expand on the history of modern computing.}

\begin{figure}[t]
    \centering
    \includegraphics[width=0.95\textwidth]{eniac_programmers.jpeg}
    \caption{Ruth Lichterman (left) and Marlyn Wescoff (middle) were two of the female programmers of the ENIAC.} %Photo: \copyright Corbis/ Getty Images.}
    \label{fig:eniac_programmers}
\end{figure}

With the advent of the internet and the widespread adoption of digital technologies, the amount of data available to decision-makers has grown exponentially. 
This has led to a surge of interest in the scientific field of machine learning, a subfield of artificial intelligence that focuses on developing algorithms that can learn from data to recognize salient patterns and make predictions to guide decisions. 
Enabled by a massive increase in the power and availability of parallel computing resources, such methods have set new standards for the performance of computer systems in various domains, including speech recognition, computer vision, and natural language processing \cite{lecun_deep_2015}. 
In recent years, machine learning has also been applied within healthcare, where it has shown promising results in tasks such as medical imaging \cite{lundervold_overview_2019}, drug discovery \cite{chen_rise_2018}, and clinical decision support \cite{cite15, cite14}.

However, these technologies come with the same risks as previous generations did, and in some cases, they may exacerbate them. 
Inherent limitations of the technology itself, human error, and the complexity of the real world are all factors that can lead to failures. A pertinent example is the nascent use of machine learning in autonomous driving systems: 
In 2018, the first incident of a self-driving car killing a pedestrian took place in Tempe, Arizona. 
The car, an SUV operated by Uber, was driving in autonomous mode when it struck a pedestrian crossing the street with her bicycle. 
According to the investigation by the National Transportation Safety Board, the car's sensors detected the pedestrian $5.6$ seconds before the crash, but it did not correctly predict her path or reduce the SUV's speed (\cref{fig:uber_nhsa_accident}). 
Specifically, during those seconds, the system incorrectly classified the pedestrian more than ten times, first as a vehicle, then as an unknown object and ultimately as a bicyclist, each time changing, or resetting, the predicted path. 
About one second before the crash, the system determined that a collision was imminent, but the situation exceeded the constraints within which the autonomous driving system was allowed to operate, and the car's safety driver failed to intervene \cite{nationaltransportationsafetyboardnhsa_collision_2019}.

@@

Though this incident is an extreme one, it highlights the challenges of applying machine learning in real-world scenarios and underlines the need for machine learning systems to be robust to adverse or difficult situations.
Recently, European policymakers proposed a set of requirements for applications of machine learning to ensure that they are safe, reliable, and trustworthy \cite{europeancommission_briefing_2021}. This AI Act groups applications into three categories based on the potential risk they pose to society: limited risk, high risk, and unacceptable risk. Limited risk applications, for instance within gaming, must meet transparency requirements, and unacceptable risk applications, such as public use of facial recognition, are banned. High risk applications include those in healthcare and must meet a strict set of requirements focusing on anti-discrimination and robustness. 
%
One of the key ways to improve the safety, reliability and trustworthiness of machine learning systems is to equip them with the ability accurately estimate the uncertainty of their predictions. 
A model that can accurately estimate the uncertainty of its predictions can communicate to other systems or humans when it is unsure about a prediction, and so avoid making potentially catastrophic mistakes.
Fittingly, a recent study by the European Parliamentary Research Services concluded:

\begin{center}

\textit{"Future AI solutions for healthcare should be implemented by integrating uncertainty estimation, a relatively new field of research that aims to provide clinicians with clinically useful indications on the degree of confidence in AI predictions"} \cite{europeanparliament_artificial_2022}.

\end{center}

@@

% OUTLINE: Primer overview of the thesis. What are the topics and what is the context of the research.
\noindent The research presented in this thesis will focus on the use of machine learning in relation to uncertainty estimation, speech processing, and medical conversations. 
The research project of this thesis was formulated and carried out in collaboration with the Danish company Corti, which develops computer software for medical decision support. 
In the rest of this chapter, we will provide a high-level introduction to these topics and discuss the motivation behind the research project.

\section{Motivation and Corti's use-cases}
%
% OUTLINE: Motivate the research project. 
% - What is the context of the research and why is it important? 
% - What are the current applications of machine learning in the domain (decision-support in healthcare)?
% - What are the practical implications of the research? 
% - What are the challenges and limitations of the current state-of-the-art?
%
At the start of this project in 2020, Corti's main product was a software system providing decision-support for emergency medical dispatchers which operate the 1-1-2 emergency line in Copenhagen, Denmark and the 9-1-1 emergency line in Seattle, Washington in the United States.
The system, having been further developed, is still in use in 2023, and provides an interface for navigating a graph-based protocol, defined by the local emergency medical services, which dispatchers use to triage patients during emergency calls. 
Calls are transcribed in real-time using customer-domain automatic speech recognition and the system makes suggestions for the dispatcher to use in the conversation with the caller, including notifications about forgotten protocol questions \cite{havtorn_multiqt_2020} and potential cases of cardiac arrests \cite{cite15, cite14}. The system logs the details about the call and the dispatcher's actions to be used for later review and quality assurance.

\subsection{Stroke recognition in emergency calls}
%
In 2021, Corti entered a research collaboration with the Capital Region of Denmark and the Copenhagen Emergency Services to develop a system for recognizing stroke cases in calls to the 1-1-2 emergency line and the 1813 medical helpline. 
Stroke is one of the biggest causes of disability and death worldwide \cite{cite1,cite2,cite3} and effective treatment is highly time-sensitive \cite{cite4,cite5}. The most common gateway to specialized treatment and hospital admittance is through prehospital telehealth services like emergency medical call centers, nurse advice call lines, and out-of-hours health services \cite{cite6,cite7}, however, studies have found that approximately half of all patients with stroke do not receive the correct triage for their condition from call-takers \cite{cite10,cite11,cite12}. 
This is likely due to the complexity of stroke cases which exhibit a wide range of symptoms that can be difficult to recognize over the phone. Additionally, out of total call volume, stroke cases are relatively rare, occurring in only $0.25\%$ of all calls to the Copenhagen 1-1-2 emergency line in 2021. This makes it difficult for call-takers to gain practical experience in recognizing stroke cases which exacerbates the general difficulty of recognizing the cases in the first place. Although several efforts have been made to improve the recognition of stroke cases in emergency calls \cite{cite13,cite14,cite15}, there is still a need for better tools to support call-takers in recognizing stroke cases.

A reasonable machine learning pipeline to assist in recognizing stroke on calls to emergency services might consist of an automatic speech recognition model to transcribe the conversation and a binary classifier to estimate the probability that what the transcript describes is a case of stroke. 
Such a system might be trained on a dataset calls with verified stroke and non-stroke cases and evaluated against a held-out test set of calls where the call-taker has indicated that they suspect a stroke. Due to the low prevalence of stroke cases, the dataset would likely be imbalanced, and the number of stroke cases limited. 
% The performance of such a system would depend on the quality of the automatic speech recognition model, the quality of the classifier, and the quality of the data used to train them. 
In such a system, many factors can lead to classification errors, and when that happens, we would like the model to express high degree of uncertainty. 
The automatic speech recognition model might be make an error in the transcription of particular word or phrase due to a noisy environment, overlapping, slurred or mumbling speech, or words not in the model's training data. 
The classifier might misclassify a conversation due to medically ambiguous symptoms, a general lack of information in the conversation, or errors of the speech recognizer propagated via the transcription. 
How to best make such models accurately represent and estimate the uncertainty of given predictions, and how to use it to improve the value of such a system, especially in cooperation with the call taker, is an open question.
What is clear, however, is that the ability of such a system to accurately estimate the uncertainty of given predictions is crucial for its safe and reliable deployment in the real world. Underconfident predictions could lead to unnecessary delays in treatment and increased costs for the healthcare system. Overconfident predictions could lead to misdiagnoses and potentially fatal consequences for patients.

\subsection{Medical coding of clinical notes}
%
During the course of the project presented in this thesis, Corti's product portfolio has expanded to include several new products. 
One of these is a software system for medical coding of clinical notes. 
When a patient is admitted to a hospital, the medical staff will write a clinical note describing the patient's diagnosis and the procedures they underwent, which includes drug prescriptions. 
These notes are then used for billing and reimbursement purposes, as well as for research and quality assurance. 
The clinical notes are written in natural language adhering to a certain structure, usually by a doctor. Later, a medical coder will assign a set of medical codes to the note based on its content. 
The process of medical coding is time-consuming and error-prone due to the number and complexity of medical diagnoses and procedures. For instance, the widely used International Classification of Diseases (ICD) standard consist of 55,000 medical codes in version 10 and 85,000 in version 11 \cite{worldhealthorganisationwho_international_2023}. Additionally, a single medical note will usually contain several diagnoses and procedures which must all be inferred from the clinical note. In public benchmarking datasets such as MIMIC-III and IV the number is around 15. Official guidelines describe how to code specific diagnoses and procedures and combinations thereof \cite{centersformedicaremedicaidservicesus_icd10cm_2023}. 
For instance, a code may have several exclusion or inclusion criteria which define when certain other codes must be coded along with it, or when they must not.

\todo[inline]{
Discuss the role of uncertainty estimation in medical coding. 
Special challenges include significant class imbalance that cannot be removed by stratification since each example is associated with multiple labels.
Mutually dependent classes also complicates uncertainty estimation. Some classes are mutually exclusive, while others are not. Some classes are easily confused with each other. Uncertainty can be helpful in identifying code candidates as uncertain codes may need human review.
}

Many codes are rarely used and datasets for medical coding usually have a considerable class-imbalance which can be impossible to exactly correct for by stratified sampling since each example is associated with multiple labels.

\todo[inline]{Write paragraph on using machine learning for medical coding and the role of uncertainty estimation in that.}

\section{Machine learning reliability}
%
Many modern machine learning models are deep neural networks with millions or even billions of parameters. 
These models are often referred to as black-box models as it is difficult to understand how they arrive at their predictions. 
The lack of interpretability makes it difficult to assess the reliability of the model's predictions. 
Furthermore, a number of factors, including the ability of deep neural networks to overfit, or even memorize, the training data \cite{arpit_closer_2017, burg_memorization_2021}, and the use of training objectives that are proxies for the evaluation metrics of interest, lead to models that often are miscalibrated; that is, the probability of a class predicted by the model does not reflect the true probability of the predicted class under the circumstances \cite{guo_calibration_2017, kull_temperature_2019}. Usually, the predicted probabilities are too extreme which leads to overconfident predictions. This means a model will often assign a high probability to the predicted class, even when it is wrong. 
In practice, such a model can be hard to trust. No model is perfect, but the inability to judge the certainty of a model's predictions makes it difficult to know how to weigh the model's predictions in a decision-making process.

A connection can be made back to the incident with the Uber self-driving car. 
The multitude of rapid misclassifications made by the Uber self-driving car in Tempe is likely an indication that the model responsible for object-classification was overconfident in its predictions and unable to represent or communicate uncertainty to other systems in the car, or the safety driver. Instead, the model's predictions were taken at face value and subsystems in the car acted on them without appropriate consideration to their reliability.

\subsection{Model calibration}

\todo[inline]{
    Expand in this discussion of overconfidence and calibration. E.g.: Calibration is fairly well understood in the context of classification, especially binary, but is harder in structured prediction such as automated speech recognition. Also: 
    Even if a model is well-calibrated, it can still be overconfident in wrong predictions - this is a problem in particular for rare events or adverse examples that were not presented to the model during training. Such examples are sometimes called out-of-distribution examples with reference to the training distribution. 
    For instance, a model trained to classify images of cats and dogs may have been well-calibrated, but if it is presented with an image of a horse, it has no option but to assign 100\% total probability to the cat and dog categories, even though that is clearly wrong. 
    Worse yet, we might expect the model to then at least assign 50\% probability to each of the cat and dog categories, but that behavior is not guaranteed. Specifically, since no horses were in the training data, the model will assign arbitrary probabilities to either class and so still risks being confidently wrong.
}

\subsection{Understanding uncertainty}
% 
% OUTLINE: Introduce aleatoric and epistemic uncertainty. 
% - Discuss the distinction between them and how they relate to the problem of uncertainty estimation in machine learning.
% - 
% 
%
The distinction in uncertainty estimation between known and unknown unknowns is sometimes referred to in terms of \textit{aleatoric uncertainty} and \textit{epistemic uncertainty} \cite{kendall_what_2017}. Both categories of uncertainty are usually used to describe uncertainties in a model's predictions or the model itself including its parameters, but they differ in their sources. 
Aleatoric uncertainty (from latin (aleatorius) 'dice player' or 'gambler') refers to uncertainty in the data itself due to true randomness which is irreducible even with infinite data. 
Epistemic uncertainty (from Greek (epistḗmē) 'knowledge') refers to uncertainty due to things one could in principle know but does not in practice. 
% It could arise because a measurement is not accurate, because the model neglects certain effects, or because particular data have been deliberately hidden. 
Epistemic uncertainty is sometimes referred to as \textit{systematic uncertainty} and aleatoric uncertainty as \textit{stochastic uncertainty} \cite{kendall_what_2017}. 
%, or, \textit{reducible} and \textit{irreducible} uncertainty, respectively. This latter terminology is somewhat misleading since epistemic uncertainty might not be reducible by additional data collection if the model is not sufficiently expressive to represent the data distribution, and aleatoric uncertainty is not necessarily irreducible if we can reduce inherent noise by improving the instruments used to collect the data \cite{kendall_what_2017}. 
Although the distinction between aleatoric and epistemic uncertainty is important, it is not always useful. For instance, a model may be uncertain about a particular prediction due to any of these two kinds of uncertainty. For that reason, the quantity of interest is usually the ``total" uncertainty, often referred to as \textit{predictive uncertainty}, which is simply the combined epistemic and aleatoric uncertainties.

An example of aleatoric uncertainty is the uncertainty in the transcription of a word due to a noisy environment, overlapping, slurred or mumbling speech. We can reduce this uncertainty by improving the automatic speech recognition model, but we cannot eliminate it completely.

Words not present in the training data are arguably sources of epistemic uncertainty although a good speech recognition model may generalize well to unseen words if its spelling and pronunciation follow the same patterns as the words in the training data.

A stronger example of epistemic uncertainty is the occurrence of truly out-of-distribution examples, such as the horse in the image classifier example above: inputs unlike anything the model has seen during training that potentially require outputs unavailable to the model. 
In high-dimensional data with a practically unlimited diversity in possible out-of-distribution examples, it is impossible to collect enough data to eliminate epistemic uncertainty. If we included some horses in the training data, the model would likely be able to recognize them, but it would still be unable to recognize giraffes, or zebras, or unicorns. 
Ultimately, any practical machine learning model will have at least some sources of epistemic uncertainty i.e. unknown sources of uncertainty.

% Aleatory uncertainty is often pretty straightforward to model since it is available in the data. Epistemic is much harder since we don't know much about it, other than we have only examples of where it is low.

% In real life applications, both kinds of uncertainties are present. Uncertainty quantification intends to explicitly express both types of uncertainty separately. The quantification for the aleatoric uncertainties can be relatively straightforward, where traditional (frequentist) probability is the most basic form. Techniques such as the Monte Carlo method are frequently used. A probability distribution can be represented by its moments (in the Gaussian case, the mean and covariance suffice, although, in general, even knowledge of all moments to arbitrarily high order still does not specify the distribution function uniquely), or more recently, by techniques such as Karhunen–Loève and polynomial chaos expansions. To evaluate epistemic uncertainties, the efforts are made to understand the (lack of) knowledge of the system, process or mechanism. Epistemic uncertainty is generally understood through the lens of Bayesian probability, where probabilities are interpreted as indicating how certain a rational person could be regarding a specific claim.

% In mathematics, uncertainty is often characterized in terms of a probability distribution. From that perspective, epistemic uncertainty means not being certain what the relevant probability distribution is, and aleatoric uncertainty means not being certain what a random sample drawn from a probability distribution will be.

\todo[inline]{Wrap this up. How to model aleatoric and epistemic uncertainty. Why do we care most about epistemic uncertainty (because we care about correctly estimating predictive uncertainty, and aleatoric is easy to model).}

\section{Scope of the thesis}
%
% OUTLINE: 
% - Tie together the different topics of the thesis and point briefly to the relevant papers.
%
The complete set of research published as part of this project covers a number of topics related to decision-support in medical conversations. 
One group of the produced studies is concerned with generative latent variable models and their applications to uncertainty estimation and speech modelling \cite{havtorn_hierarchical_2021,havtorn_benchmarking_2022,bergamin_modelagnostic_2022}. 
Another group of studies is concerned with self-supervised speech representation learning and automatic speech recognition \cite{borgholt_scaling_2021,borgholt_we_2021,mohamed_selfsupervised_2022,borgholt_brief_2022}. 
A final group deals with applications of machine learning within the medical domain, including recognition of stroke cases in calls to medical helplines \cite{wenstrup_retrospective_2023} and medical coding of clinical notes \cite{edin_automated_2023}.

@@
